---
title: "STA 322 Project 1: Sampling Design"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Frame

Sampling frame: 36 Trinity units (departmental websites), N = total faculty listed.
ROTC units and Health, Wellness & PE are kept in the frame.

```{r frame}
frame <- read.csv("faculty_counts.csv", stringsAsFactors = FALSE)
frame$non_phd_count <- frame$total_faculty - frame$phd_count
N <- sum(frame$total_faculty)
n <- 100
c(N = N, n = n, sampling_fraction = n / N)
```

## Strata

Stratified sampling with allocation proportional to size (PPS allocation), which
gives every faculty member (nearly) the same inclusion probability n/N, so the design
is approximately self-weighting. To keep the weights equal, no stratum may need a
minimum sample size override: every stratum must have a proportional share
n * N_h / N of at least 2 (so N_h >= about 17.3), which also gives a within-stratum
variance estimate. Departments below that size are pooled with related units. Every
other department is its own stratum.

```{r strata}
pool <- list(
  "Military & PE (pooled)" = c("Air Force ROTC", "Army ROTC", "Navy ROTC",
                               "Health, Wellness & Physical Education"),
  "Small languages (pooled)" = c("International Comparative Studies",
                                 "Slavic & Eurasian Studies", "Romance Studies",
                                 "German Studies"),
  "Dance & Theater (pooled)" = c("Dance", "Theater Studies"),
  "Education & Gender Studies (pooled)" = c("Education",
                                            "Gender, Sexuality & Feminist Studies"),
  "Classical & Religious Studies (pooled)" = c("Classical Studies",
                                               "Religious Studies")
)
frame$stratum <- frame$department
for (s in names(pool)) frame$stratum[frame$department %in% pool[[s]]] <- s

strata <- aggregate(cbind(N_h = total_faculty, phd_h = phd_count) ~ stratum,
                    data = frame, FUN = sum)
nrow(strata)
```

## Allocation

Proportional allocation with largest-remainder rounding so the total is exactly
n = 100. Because n_h must be an integer, weights N_h / n_h are equal only up to
rounding; the design weight is still used in estimation.

```{r alloc}
strata$exact <- n * strata$N_h / N
strata$n_h <- floor(strata$exact)
short <- n - sum(strata$n_h)
bump <- order(strata$exact - strata$n_h, decreasing = TRUE)[seq_len(short)]
strata$n_h[bump] <- strata$n_h[bump] + 1
strata$weight <- strata$N_h / strata$n_h   # design weight, ~N/n
stopifnot(sum(strata$n_h) == n, all(strata$n_h >= 2))
knitr::kable(strata[order(strata$stratum), ], digits = 2,
             col.names = c("Stratum", "N_h", "PhDs", "Exact n_h", "n_h", "Weight"))
range(strata$weight)
```

## Draw the sample

Random positions within each stratum's roster, with a fixed seed. For the
pooled strata, positions run through the member departments in the order below,
then are mapped back to a department and a position within that department's
roster list (order on the department website).

```{r draw}
set.seed(322)
rows <- list()
for (i in seq_len(nrow(strata))) {
  s <- strata$stratum[i]
  members <- frame[frame$stratum == s, ]
  pos <- sort(sample(strata$N_h[i], strata$n_h[i]))
  ends <- cumsum(members$total_faculty)
  starts <- ends - members$total_faculty
  idx <- findInterval(pos - 1, ends) + 1
  rows[[i]] <- data.frame(stratum = s, department = members$department[idx],
                          roster_position = pos - starts[idx])
}
sample_ids <- do.call(rbind, rows)
sample_ids$weight <- strata$weight[match(sample_ids$stratum, strata$stratum)]
nrow(sample_ids)
write.csv(sample_ids, "sample_ids.csv", row.names = FALSE)
knitr::kable(sample_ids[, c("department", "roster_position")], row.names = FALSE)
```

The sample size per department:

```{r bydept}
knitr::kable(as.data.frame(table(Department = sample_ids$department)),
             col.names = c("Department", "Sample size"))
```

## Survey design object

Stratified SRS without replacement: strata are the 30 strata above, `fpc` is the
stratum population size N_h, and the weight is N_h / n_h.

```{r design-check, message=FALSE}
library(survey)
sample_ids$N_h <- strata$N_h[match(sample_ids$stratum, strata$stratum)]
sample_ids$one <- 1
svytotal(~one, svydesign(ids = ~1, strata = ~stratum, weights = ~weight,
                         fpc = ~N_h, data = sample_ids))
```

## Estimation (run after data collection)

Fill in `faculty_sample.csv` with one row per sampled faculty member: the columns of
`sample_ids.csv` plus `female` (0/1), `ongoing_grant_2026` (0/1), `has_phd` (0/1),
`phd_year` (blank if no PhD) and any Q4 variable. This section runs only once the file exists.

```{r estimate, eval=file.exists("faculty_sample.csv"), message=FALSE}
dat <- read.csv("faculty_sample.csv")
dat$N_h <- strata$N_h[match(dat$stratum, strata$stratum)]
dat$weight <- strata$weight[match(dat$stratum, strata$stratum)]
dat$years_since_phd <- ifelse(dat$has_phd == 1, 2026 - dat$phd_year, NA)
des <- svydesign(ids = ~1, strata = ~stratum, weights = ~weight,
                 fpc = ~N_h, data = dat)

# Q1 and Q2: proportions with 95% CIs
q1 <- svyciprop(~female, des);             q1; confint(q1)
q2 <- svyciprop(~ongoing_grant_2026, des); q2; confint(q2)

# Q3: mean among PhD holders only. subset() on the design object is the correct
# domain estimator (a ratio estimate), not re-analyzing a separate sample.
des_phd <- subset(des, has_phd == 1)
q3 <- svymean(~years_since_phd, des_phd); q3; confint(q3)
```
