---
title: "STA 322 Project 1: Sampling Design"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Frame

Sampling frame: 36 Trinity units (departmental websites), N = total faculty listed.
ROTC units and Health, Wellness & PE are kept in the frame.

```{r frame}
frame <- read.csv("faculty_counts.csv", stringsAsFactors = FALSE)
frame$non_phd_count <- frame$total_faculty - frame$phd_count
N <- sum(frame$total_faculty)
n <- 100
c(N = N, n = n, sampling_fraction = n / N)
```

## Strata

Stratified sampling with allocation proportional to size (PPS allocation), which
gives every faculty member (nearly) the same inclusion probability n/N, so the design
is approximately self-weighting. To keep the weights equal, no stratum may need a
minimum sample size override: every stratum must have a proportional share
n * N_h / N of at least 2 (so N_h >= about 17.3), which also gives a within-stratum
variance estimate. Departments below that size are pooled with related units. Every
other department is its own stratum.

```{r strata}
pool <- list(
  "Military & PE (pooled)" = c("Air Force ROTC", "Army ROTC", "Navy ROTC",
                               "Health, Wellness & Physical Education"),
  "Small languages (pooled)" = c("International Comparative Studies",
                                 "Slavic & Eurasian Studies", "Romance Studies",
                                 "German Studies"),
  "Dance & Theater (pooled)" = c("Dance", "Theater Studies"),
  "Education & Gender Studies (pooled)" = c("Education",
                                            "Gender, Sexuality & Feminist Studies"),
  "Classical & Religious Studies (pooled)" = c("Classical Studies",
                                               "Religious Studies")
)
frame$stratum <- frame$department
for (s in names(pool)) frame$stratum[frame$department %in% pool[[s]]] <- s

strata <- aggregate(cbind(N_h = total_faculty, phd_h = phd_count) ~ stratum,
                    data = frame, FUN = sum)
nrow(strata)
```

## Allocation

Proportional allocation with largest-remainder rounding so the total is exactly
n = 100. Because n_h must be an integer, weights N_h / n_h are equal only up to
rounding; the design weight is still used in estimation.

```{r alloc}
strata$exact <- n * strata$N_h / N
strata$n_h <- floor(strata$exact)
short <- n - sum(strata$n_h)
bump <- order(strata$exact - strata$n_h, decreasing = TRUE)[seq_len(short)]
strata$n_h[bump] <- strata$n_h[bump] + 1
strata$weight <- strata$N_h / strata$n_h   # design weight, ~N/n
stopifnot(sum(strata$n_h) == n, all(strata$n_h >= 2))
knitr::kable(strata[order(strata$stratum), ], digits = 2,
             col.names = c("Stratum", "N_h", "PhDs", "Exact n_h", "n_h", "Weight"))
range(strata$weight)
```

## Draw the sample

Random positions within each stratum's roster, with a fixed seed. For the
pooled strata, positions run through the member departments in the order below,
then are mapped back to a department and a position within that department's
roster list (order on the department website).

```{r draw}
set.seed(322)
rows <- list()
for (i in seq_len(nrow(strata))) {
  s <- strata$stratum[i]
  members <- frame[frame$stratum == s, ]
  pos <- sort(sample(strata$N_h[i], strata$n_h[i]))
  ends <- cumsum(members$total_faculty)
  starts <- ends - members$total_faculty
  idx <- findInterval(pos - 1, ends) + 1
  rows[[i]] <- data.frame(stratum = s, department = members$department[idx],
                          roster_position = pos - starts[idx])
}
sample_ids <- do.call(rbind, rows)
sample_ids$weight <- strata$weight[match(sample_ids$stratum, strata$stratum)]
nrow(sample_ids)
write.csv(sample_ids, "sample_ids.csv", row.names = FALSE)
knitr::kable(sample_ids[, c("department", "roster_position")], row.names = FALSE)
```

The sample size per department:

```{r bydept}
knitr::kable(as.data.frame(table(Department = sample_ids$department)),
             col.names = c("Department", "Sample size"))
```

## Survey design object

Stratified SRS without replacement: strata are the 30 strata above, `fpc` is the
stratum population size N_h, and the weight is N_h / n_h.

```{r design-check, message=FALSE}
library(survey)
sample_ids$N_h <- strata$N_h[match(sample_ids$stratum, strata$stratum)]
sample_ids$one <- 1
svytotal(~one, svydesign(ids = ~1, strata = ~stratum, weights = ~weight,
                         fpc = ~N_h, data = sample_ids))
```

## Estimation

Reads the collected data (`data.csv`). Emeriti are kept as drawn, since dropping them
would change the selection probabilities. Missing values (a few faculty with no
education or grant information on Scholars, or an unstated sex) are handled by
estimating over the faculty with an observed value, using `subset()` on the design
object so the variance still reflects the full stratified design.

```{r estimate, message=FALSE}
dat <- read.csv("data.csv", na.strings = c("NA", ""), stringsAsFactors = FALSE)
stopifnot(all(dat$stratum %in% strata$stratum))
dat$N_h <- strata$N_h[match(dat$stratum, strata$stratum)]
dat$weight <- strata$weight[match(dat$stratum, strata$stratum)]
dat$years_since_phd <- ifelse(dat$has_phd %in% 1, 2026 - dat$phd_year, NA)
dat$sex <- factor(dat$female, levels = c(0, 1), labels = c("Men", "Women"))
des <- svydesign(ids = ~1, strata = ~stratum, weights = ~weight,
                 fpc = ~N_h, data = dat)

# Missing-data summary for the variables used
colSums(is.na(dat[, c("female", "ongoing_grant_2026", "has_phd", "phd_year",
                      "full_professor")]))
```

### Q1. Fraction of faculty who are women

```{r q1}
d1 <- subset(des, !is.na(female))
q1 <- svyciprop(~female, d1)
q1
```

### Q2. Fraction with at least one grant ongoing in 2026

```{r q2}
d2 <- subset(des, !is.na(ongoing_grant_2026))
q2 <- svyciprop(~ongoing_grant_2026, d2)
q2
```

### Q3. Average years since PhD among PhD holders

`subset()` on the design gives the domain (ratio) estimator for the mean among
PhD holders. Faculty with a PhD but no year listed are dropped from this estimate.

```{r q3}
d3 <- subset(des, has_phd %in% 1 & !is.na(years_since_phd))
q3 <- svymean(~years_since_phd, d3)
q3

# Estimated fraction of faculty holding a PhD (for context)
q3b <- svyciprop(~I(has_phd == 1), subset(des, !is.na(has_phd)))
q3b
```

### Q4. Are women less likely than men to be full professors?

`full_professor` is 1 for rank "Professor" (including Professor of the Practice,
distinguished professors, and emeriti with that rank) and 0 otherwise.

```{r q4}
d4 <- subset(des, !is.na(female) & !is.na(full_professor))
# fraction full professors overall and by sex
q4_all <- svyciprop(~full_professor, d4)
q4_all

q4_by <- svyby(~full_professor, ~sex, d4, svymean, vartype = c("se", "ci"),
                covmat = TRUE)
q4_by
# difference (women - men) and a design-based test of association
svycontrast(q4_by, c(-1, 1))
svychisq(~full_professor + sex, d4)
```

### Summary table

```{r summary}
est <- function(x, ci) c(Estimate = as.numeric(x), Lower = ci[1], Upper = ci[2])
res <- rbind(
  "Q1 Women" = est(q1, confint(q1)),
  "Q2 Ongoing grant in 2026" = est(q2, confint(q2)),
  "Q3 Years since PhD (PhD holders)" = est(q3, confint(q3)),
  "Q4 Full professor (all)" = est(q4_all, confint(q4_all)),
  "Q4 Full professor, women" = unlist(q4_by[2, c("full_professor", "ci_l", "ci_u")]),
  "Q4 Full professor, men" = unlist(q4_by[1, c("full_professor", "ci_l", "ci_u")])
)
colnames(res) <- c("Estimate", "95% CI lower", "95% CI upper")
knitr::kable(res, digits = 3)
```
